A Comparison of Parallel Graph Processing Benchmarks

### 2017-02-17 papers/publication/paper-pollard.tex
fixed num cores

At the human level, there are optimizations for each system which may not be apparent. For example, the Parallel Boost Graph Library (PBGL) \cite{Gregor:2005:PBGL} provides generic implementations of their algorithms and the programmer must provide the template specializations. The optimal data structures may differ across algorithms and graphs and must be determined by the programmer.

@@

\subsection{Machine Specifications}
Table~\ref{tab:specs} shows the specifications of the server on which we performed all experiments.

@@

\begin{table}
	\centering
	\caption{The operating system is GNU/Linux version 4.4.0-22. The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit. We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.}
	% For arya I deleted Max RAM Freq	2133MHz
	\pgfplotstabletypeset[
	header=false,
	col sep=tab,
	string type,
	every head row/.style={output empty row, before row=\bottomrule},
	columns/0/.style={column type={|p{1in}|}},
	columns/1/.style={column type={p{3.2in}|}},
	every last row/.style={after row=\toprule},
	]{../report/specs.csv}
	\label{tab:specs}
\end{table}

@@

We attempt to use similar stopping criteria for each system, but GraphMat executes until no vertices change rank; effectively its stopping criterion requires the $\infty$-norm be less than machine epsilon. This could account for the increased number of iterations. We adjusted the other systems to use $\sum_{k=1}^{n} |p_k^{(i)} - p_k^{(i-1)}| < \epsilon $ as the stopping criteria, where $i$ is the iteration and $n$ is the number of vertices. We use $\epsilon = 6 \times 10^{-8}$ because this value is approximately machine epsilon for a single precision floating-point number. The goal is to make the stopping criteria for all implementations as similar as possible.

@@

The challenge in comparing iteration counts for PageRank underscores an important challenge for any comparison of graph processing systems. The assumptions under which the various platforms operate can have a dramatic effect on the program. For example, the GAP Benchmark Suite stores vertex weights as 32-bit integers. However, other systems store them as floating point numbers. This may affect performance in addition to runtime behavior in cases where weights like $0.2$ are cast to $0$. Similarly, how a graph is represented in the system (e.g. weighed or directed) may have performance and algorithmic implications but is not always readily apparent.

Figure~\ref{fig:bfs-scaling} shows the scalability and speedup for BFS. We define the strong scaling as $T_1 / (n T_n)$ where $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ is the time for $n$ threads. A linear scalability is when $T_n = T_1/n$ and is the horizontal line on the bottom plot in Fig.~\ref{fig:bfs-scaling}. These plots show in general poor scaling throughout, though GraphBIG has the best scaling. A potential explanation is GraphBIG is the slowest system so there is more room for improvement. Additionally, $2^{20}$ vertices is a ``small'' graph by today's standards and thus library designers focus on scalability for larger graphs. Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine.
\begin{figure}
	\centering
	\begin{minipage}{0.9\linewidth}
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup.pdf}
	\end{minipage}
	\begin{minipage}{0.9\linewidth}
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss.pdf}
	\end{minipage}
	\caption{The top figure has both axes logarithmic with an exception at 72 threads for readability. Linear speedup is. GraphMat dips below 1 because it is slower for $2$ threads than for 1. $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time for $n$ threads.}
	\label{fig:bfs-scaling}
\end{figure}
\todo{Rerun GraphMat on scale 22 with 2 threads. Surprising that it slows down}

>>>>

From a graph algorithm user's point of view, there are optimizations for each system which may not be apparent. For example, the Parallel Boost Graph Library (PBGL)~\cite{Gregor:2005:PBGL} provides generic implementations of their algorithms and the programmer must provide the template specializations. The optimal data structures may differ across algorithms and graphs and must be determined by the programmer.

@@

\subsection{Machine Specifications}
We performed experiments on our 32-core Intel Haswell server, whose specifications are outlined below.
%Table~\ref{tab:specs} shows the specifications of the server on which we performed all experiments.
%\begin{table}[h!tb]
%	\centering
%	\caption{
The operating system is GNU/Linux version 4.4.0-22. The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit. We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.
\begin{center}
	% For arya I deleted Max RAM Freq	2133MHz
	\pgfplotstabletypeset[
	header=false,
	col sep=tab,
	string type,
	every head row/.style={output empty row, before row=\bottomrule},
	columns/0/.style={column type={|p{1in}|}},
	columns/1/.style={column type={p{3.2in}|}},
	every last row/.style={after row=\toprule},
	]{../report/specs.csv}
	%\label{tab:specs}
\end{center}

@@

@@

We attempt to define similar stopping criteria for each system, but GraphMat executes until no vertices change rank; effectively its stopping criterion requires the $\infty$-norm be less than machine epsilon. This could account for the increased number of iterations. We adjusted the other systems to use $\sum_{k=1}^{n} |p_k^{(i)} - p_k^{(i-1)}| < \epsilon $ as the stopping criteria, where $i$ is the iteration and $n$ is the number of vertices. We use $\epsilon = 6 \times 10^{-8}$ because this value is approximately machine epsilon for a single precision floating-point number. The goal is to make the stopping criteria for all implementations as similar as possible.

@@

The difficulty in comparing iteration counts for PageRank underscores an important challenge for any comparison of graph processing systems. The assumptions under which the various platforms operate can have a dramatic effect on the program. For example, the GAP Benchmark Suite stores vertex weights as 32-bit integers. However, other systems store them as floating-point numbers. This may affect performance in addition to runtime behavior in cases where weights like $0.2$ are cast to $0$. Similarly, how a graph is represented in the system (e.g., weighed or directed) may have performance and algorithmic implications but is not always readily apparent.

Figure~\ref{fig:bfs-scaling} shows the scalability and speedup for BFS. We define the strong scaling as $T_1 / (n T_n)$ where $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ is the time for $n$ threads. A linear scalability is when $T_n = T_1/n$ and is the horizontal line on the bottom plot in Fig.~\ref{fig:bfs-scaling}. These plots show in general poor scaling throughout, although GraphBIG has the best scaling. A potential explanation is GraphBIG is the slowest system, so there is more room for improvement. Additionally, $2^{20}$ vertices is a ``small'' graph by today's standards and thus library designers focus on scalability for larger graphs. Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine.
\begin{figure}
	\centering
	\begin{minipage}{0.9\linewidth}
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup.pdf}
	\end{minipage}
	\begin{minipage}{0.9\linewidth}
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss.pdf}
	\end{minipage}
	\caption{The top figure has both axes logarithmic with an exception at 72 threads for readability. Linear speedup is. GraphMat dips below 1 because it is slower for $2$ threads than for 1. $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time for $n$ threads.}
	\label{fig:bfs-scaling}
\end{figure}
\todo{Rerun GraphMat on scale 22 with 2 threads. Surprising that it slows down}
